Confirmation of Wide-Field Signatures in Redshifted 21~cm Power Spectra using Murchison Widefield Array Observations

### 2015-04-07 ms.tex
Modify Introduction, add new references to eor.bib
From af1b5c4:
============

In ms.tex,

(*) Modify Introduction

In eor.bib,

(*) Add references to Di Matteo et al. 2004, Gnedin & Shaver 2004,
Santos et al. 2005, Wang et al. 2006, and Gleser et al. 2008.

While sensitive instruments such as the Square Kilometre Array (SKA) with the capability of direct imaging of redshifted H~{\sc i} become operational, numerous pathfinders and precursors such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13,bow13}, the Low Frequency Array \citep[LOFAR;][]{van13}, the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10}, and the planned Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} (HERA) have enough sensitivity for a statistical detection of the signal \citep{bow06,par12a,bea13,dil13,thy13,pob14} and most of these are already operational.

The primary challenge to detection of cosmological H~{\sc i} from the EoR comes from continuum emission from Galactic and extragalactic foreground objects, relative to which the desired signal is $\sim 10^4$ times weaker \citep[see, e.g.,][]{dim02,zal04,fur06,ali08,ber09,ber10,gho12}. Since the foregrounds and the signal have inherent differences in spatial isotropy and spectral smoothness \citep{mor04,mor06,bow09,liu11,par12b,dil13,pob13}, a detailed characterization of foreground emission has become essential \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,dil14,liu14a,liu14b,thy13,thy15}.

A recent study by \citet{thy15} using instrument and all-sky foreground models has shed new light on the significant role played by the interplay between foreground emission and the nature of wide-field measurements, typical of most modern EoR experiments. It was noted to cause significant contamination in Fourier modes considered to be most sensitive for the cosmological H~{\sc i} signal. It has been shown that a careful design of antenna aperture can significantly mitigate this contamination. And an optimal weighting of foreground-contaminated Fourier modes may be required to extract the signal with maximum sensitivity. Thus, knowledge of such detailed foreground signatures becomes essential for design and analysis choices of future instruments such as HERA and SKA. This paper presents a follow-up analysis of that study. By using a larger volume of MWA data obtained independently over different nights pointed at the same region of sky, we confirm with high significance the presence of certain key characteristics (referred to as the {\it pitchfork} signature) of wide-field measurements predicted in the preceding study.

>>>>

Sensitive instruments such as the Square Kilometre Array (SKA) with the capability of direct imaging of redshifted H~{\sc i} are yet to become operational. In the meanwhile, the Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} (HERA) currently under construction will be much closer in its capability to detect and place definitive constraints on the reionization epoch relative to the SKA and far ahead of current instruments such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13,bow13}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10}, which have enough sensitivity for a statistical detection of the signal \citep{bow06,par12a,bea13,dil13,thy13,pob14}.

% The primary challenge to detection of cosmological H~{\sc i} from the EoR comes from continuum emission from Galactic and extragalactic foreground objects, relative to which the desired signal is $\sim 10^4$ times weaker \citep[see, e.g.,][]{dim02,zal04,fur06,ali08,ber09,ber10,gho12}. Since the foregrounds and the signal have inherent differences in spatial isotropy and spectral smoothness \citep{mor04,mor06,bow09,liu11,par12b,dil13,pob13}, a detailed characterization of foreground emission has become essential \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,dil14,liu14a,liu14b,thy13,thy15}.

The primary challenge to detection of cosmological H~{\sc i} from the EoR arises from continuum emission from Galactic and extragalactic foreground objects, relative to which the desired signal is $\sim 10^4$ times weaker. But the inherent differences in spatial isotropy and spectral smoothness can be exploited to extract the cosmological signal from foreground contamination \citep[see, e.g.,][]{dim02,dim04,zal04,fur04,mor04,san05,fur06,mcq06,mor06,wan06,gle08,ali08,ber09,ber10,gho12}. Thus, a detailed characterization of foreground emission has become essential \citep{bow09,liu09,dat10,liu11,mor12,par12b,tro12,pob13,dil13,dil14,liu14a,liu14b,thy13,thy15}.

% A recent study by \citet{thy15} using instrument and all-sky foreground models has shed new light on the significant role played by the interplay between foreground emission and the nature of wide-field measurements, typical of most modern EoR experiments. It was noted to cause significant contamination in Fourier modes considered to be most sensitive for the cosmological H~{\sc i} signal. It has been shown that a careful design of antenna aperture can significantly mitigate this contamination. And an optimal weighting of foreground-contaminated Fourier modes may be required to extract the signal with maximum sensitivity. Thus, knowledge of such detailed foreground signatures becomes essential for design and analysis choices of future instruments such as HERA and SKA. This paper presents a follow-up analysis of that study. By using a larger volume of MWA data obtained independently over different nights pointed at the same region of sky, we confirm with high significance the presence of certain key characteristics (referred to as the {\it pitchfork} signature) of wide-field measurements predicted in the preceding study.

Our recent study \citep{thy15} used all-sky foreground and instrument models for the first time in order to simulate actual EoR experiments more accurately than previous studies. Surprisingly, we found that foreground emission outside the primary beam field of view caused the most significant contamination of the cosmological H~{\sc i} power spectrum Fourier modes that are considered most sensitive for detecting the signal in delay spectrum based analyses. This contamination is the result of the interplay between foreground emission, particularly diffuse Galactic emission, and the wide-field properties typical of EoR instruments. Our simulations predicted that delay spectra from the MWA and other experiments should exhibit a characteristic ``pitchfork'' appearance with local maxima near the horizon delay limits, in addition to at the primary lobe of the power pattern.

A careful design of antenna aperture can significantly mitigate this contamination. Optimal weighting of foreground contaminated Fourier modes may be required to extract the signal with maximum sensitivity. Thus, knowledge of such detailed foreground signatures becomes essential for design and analysis choices of future instruments such as HERA and SKA.

In \citet{thy15}, we verified the general features of our simulations against MWA observations, but were unable to confirm the actual {\it pitchfork} prediction due to insufficient sensitivity in the small amount of data analyzed. Here, we use longer MWA integrations to confirm with high significance the presence of the key {\it pitchfork} characteristics of wide-field measurements predicted in the preceding study.
